---
title: "Practice Leaflet Map: Where's Whiskers?"
author: "Eric Smith"
date: "7/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### An Interactive Exploration of Fauna in the West Virginia Highlands

The <a href = 'https://www.fs.usda.gov/main/mnf/home'>Monongahela National Forest</a> in West Virginia contains unique ecological diversity. 
In this document, the *leaflet* R library is used to create an interactive hide-and-seek map game that encourages users to virtually explore and learn about the fauna of the West Virginia highlands.

To make this simple game, the clustering feature in *leaflet*'s *addMarkers()* function is used to force users to zoom in and out of map features in search of their imagined lost cat, Whiskers. 
Along the way, they will uncover other animals that live in the Forest.
These critters are randomly populated using *runif()* and are displayed using *leaflet*'s *icons()* function.
The images were produced (poorly) by the author in MS Paint.
Many of the *popup* texts include hyperlinks to websites where the reader can learn more about the Monongahela region.

### The Code

```{r Load Packages, echo=FALSE, warning=FALSE}
library(leaflet)
```

A reference data frame is defined.

This data contains:

- a reference group number
- the number of repeated instances of the animal
- a logical variable indicating whether the animal should be randomly placed on the map
- the file path for the animal's icon
- the pop-up text to display for the animal
- optional latitude and longitude if the animal will not be randomly populated

The code below shows the setup of this data frame and the first entry as an example.

```{r Define Animal Data Frame}
## Create a data frame for the different 'species' of animal

## Number of 'species'
n_group <- 10
## Latitude/ Longitude ranges
LAT_rng <- c(38.44, 39.1)
LNG_rng <- c(-80.1,-79.376)
## Define empty data frame
ref <- data.frame(group = rep(NA,n_group),
                  Num = rep(NA,n_group),
                  Rand = rep(FALSE, n_group),
                  IMG_url = rep("", n_group),
                  POP_txt = rep("",n_group),
                  latitude = rep(NA, n_group),
                  longitude = rep(NA, n_group))
## Fill by row
ref[1,] <- list(1,
                30, 
                TRUE, 
                "images/transparent_raccoon.png",
                "This is just a raccoon.",
                NA, NA)
```
```{r echo=FALSE}
ref[2,] <- list(2,
                7, 
                TRUE,
                "images/transparent_squirrel.png",
                "This is a <a href = 'http://en.wikipedia.org/wiki/Northern_flying_squirrel'> flying squirrel</a>, specifically <i> Glaucomys sabrinus fuscus</i>.",
                NA, NA)

ref[3,] <- list(3,
                4,
                TRUE,
                "images/transparent_eagle.png",
                "Look! A bald eagle!",
                NA, NA)

ref[4,] <- list(4,
                25,
                TRUE,
                "images/transparent_deer.png",
                "Oh dear. It's a deer.",
                NA, NA)

ref[5,] <- list(5,
                5,
                TRUE,
                "images/transparent_bear.png",
                "Watch out! It's a <a href = 'http://www.nps.gov/subjects/bears/safety.htm#Do'>bear</a>!",
                NA, NA)

ref[6,] <- list(6,
                7,
                TRUE,
                "images/transparent_woodcock.png",
                "You found a woodcock! What's a <a href = 'https://en.wikipedia.org/wiki/Woodcock'>woodcock</a>?",
                NA, NA)

ref[7,] <- list(7,
                3,
                TRUE,
                "images/transparent_snail.png",
                "This is a <a href ='https://www.fws.gov/northeast/pdf/Flatspired.pdf'>flat-spired three-toothed snail</a>, protected under the Endangered Species Act.",
                NA, NA)

ref[8,] <- list(8,
                1,
                FALSE,
                "images/transparent_cat.png",
                "You found Whiskers! He was in the <a href = 'http://en.wikipedia.org/wiki/Sinks_of_Gandy'>Sinks of Gandy</a>.",
                38.715,-79.643)

ref[9,] <- list(9,
                1,
                FALSE,
                "images/transparent_salamander.png",
                "WHOA! You found a <a href = 'http://en.wikipedia.org/wiki/Cheat_Mountain_salamander'>Cheat Mountain salamander</a>.",
                38.44,-79.984)

ref[10,] <- list(10,
                 1,
                 FALSE,
                 "images/transparent_falcon.png",
                 "A peregrine falcon! Maybe its nest is on <a href = 'https://en.wikipedia.org/wiki/Seneca_Rocks'>Seneca Rocks</a>.",
                 38.835, -79.376)
```

Using the information from this reference data frame, a larger data frame is produced that can be used as input for mapping with the *leaflet* package.
This larger data frame contains randomly produced locations for many of the animals.

```{r Expand Data for Mapping}
## The reference data frame above must be expanded to include repeated
## animals and randomly populated locations

## Randomly place animals
set.seed(25)
N_rand <- sum(ref$Num[ref$Rand])
df <- data.frame(group = rep(NA, N_rand),
                 latitude = runif(N_rand, LAT_rng[1], LAT_rng[2]),
                 longitude = runif(N_rand, LNG_rng[1], LNG_rng[2]))

## Add animals with specified coordinates
df <- rbind(df, ref[!(ref$Rand),c("group","latitude","longitude")])

## Assign group numbers
df$group <- do.call(c, sapply(1:n_group,
                              function(n) rep(ref$group[n],ref$Num[n])))
## Assign popup text
popups <- do.call(c, sapply(1:n_group,
                            function(n) rep(ref$POP_txt[n], ref$Num[n])))
```

Finally, *leaflet* is used to create an interactive map with 

```{r Use Leaflet to Map Data, message=FALSE}
my_map <- df %>% leaflet(width = 910, height = 562) %>%
      addTiles() %>%
      addMarkers(icon = ~ icons(iconUrl = ref$IMG_url[group]),
                 popup = popups,
                 clusterOptions = markerClusterOptions()) %>%
      addRectangles(lng1 = LNG_rng[1] - .1, lat1 = LAT_rng[1] - .1, 
                    lng2 = LNG_rng[2] + .1, lat2 = LAT_rng[2] + .1,
                    fill = FALSE,
                    color = "forestgreen",
                    weight = 4)


```

## The Game

Oh no! Your pet cat, Whiskers, has run off into the woods!
Use the map to find your cat.

*Click and drag to move. To zoom use the buttons in the top left corner or use your mouse's scroll wheel.*
*Clicking on the green and orange clusters will automatically zoom in to see what's there.*

```{r, echo=FALSE}
my_map
```
